## Crop losses


### Most Crop Damaging event in a single occurrence

In order to determine what were the most crop damaging events in a single occurrence, we need to see how damages are distributed along the occurrences.


```{r crop-distr-1, cache=TRUE}
# crop damages
rm(prop.df, prop.all.df, prop95.df) # cleannig house

crop.df <- harm.df %>% filter(!is.na(crop.ev)) %>%
                select(1:6,10) 


# quantiles
qt <- quantile(crop.df$crop.ev, probs=seq(.998,1,0.002))

# distribution plot
plt.distr.crop0 <- ggplot(crop.df, aes(log(crop.ev)))

plt.distr.crop0 <- plt.distr.crop0 + geom_density(aes(y=..scaled..)) + #xlim(0,.5) + 
        labs(title="All events", x="log(amount $)") +
        theme(plot.title = element_text(hjust = 0.5))  


# display only the qts next to harmfull events
#qt
```

<!-- Looking at this distribution, we can infer that 99% of the occurrences caused less than **`r dollar(qt[1])` in losses**. -->

On the other hand, damaging occurrences had to have damages above zero.

Now, among the damaging occurrences, we are interested in the ones whose damages are above 99% of the most common values. 



```{r crop-distr-2, cache=TRUE}

# subset for harm events
crop.df <- crop.df %>% filter(crop.ev > 0) %>%
                arrange(desc(crop.ev)) %>%
                mutate(value = dollar(crop.ev),
                        media.raw = mean(crop.ev), 
                        mediana.raw = median(crop.ev),
                        mean = dollar(media.raw),
                        median = dollar(mediana.raw),
                        rank = seq_len(length(event)))

# quantiles, same as 
# poisson.test(mean, conf.level = 0.95)

qt <- quantile(crop.df$crop.ev, probs=seq(.999,1,0.005))
#qt
```

Looking at this distribution, we can infer that **99.8% of the damaging occurrences caused up to `r dollar(qt[1])` in losses.**




```{r crop-distr-3, cache=TRUE}

# distribution plot
plt.distr.crop1 <- ggplot(crop.df, aes(log(crop.ev)))

plt.distr.crop1 <- plt.distr.crop1 + geom_density(aes(y=..scaled..)) + #xlim(0,qt[1]) + 
        labs(title="Damaging events", x="log(amount $)") +
        # scale_x_continuous(labels = dollar)+
        # theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
        theme(plot.title = element_text(hjust = 0.5))       


```


```{r crop-distr-4, cache=TRUE, fig.width=7, fig.height=3, fig.cap="Population distribution for losses / occurrences"}
grid.arrange(plt.distr.crop0, plt.distr.crop1,
             nrow=1, ncol=2)
grid.rect(gp=gpar(fill=NA))
```



In this study, we looked on the 1% most harmful occurrences. 

```{r crop-single-95, cache=TRUE}

# subset for 99% CI
crop95.df <- filter(crop.df, crop.ev>qt[1])

# create color pallete for all events
colourCount.crop.single = length(unique(crop95.df$event))
getPalette = colorRampPalette(brewer.pal(colourCount.crop.single, "Set1"))

worst.crop.single.ev <- crop95.df$event[1]
worst.crop.single.st <- crop95.df$state[1]
worst.crop.single.ct <- crop95.df$countyname[1]
worst.crop.single.dt <- crop95.df$day[1]
worst.crop.single.value <- crop95.df$value[1]
worst.crop.single.mean <- crop95.df$mean[1]
worst.crop.single.median <- crop95.df$median[1]


# print a table
kable(crop95.df[1:20,c(13,1:6,8)],
       caption=paste("Worst crops damaging occurrences, mean = ",
                    worst.crop.single.mean, " and median = ",
                    worst.crop.single.median))

```



```{r crop-single-plot, cache=TRUE, fig.cap="Worst crops damaging occurrences"}

plt.crop.single <- ggplot(crop95.df, aes(day, crop.ev, colour=event))

plt.crop.single <- plt.crop.single + geom_point() +
        geom_text(aes(label=ifelse(rank <= 3,
                as.character(day),""),
                hjust=-.03,vjust=0.5)) +
        # geom_hline(aes(yintercept = media.raw), linetype=2) +
        # geom_hline(aes(yintercept = mediana.raw), linetype=3) +
        labs(title="Crop Dammaging",
                    y="", x="") + 
                
        expand_limits(x=as.Date('2017-01-01'))+ #ok
        scale_y_continuous(labels = dollar)+
        
        scale_colour_manual(values = getPalette(colourCount.crop.single))+                
        theme(legend.title=element_blank()) +
        theme(legend.position="bottom") +
        guides(fill=guide_legend(nrow=5, byrow=TRUE)) +
        theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
        theme(plot.title = element_text(hjust = 0.5)) 

plt.crop.single + labs(title="Worst crops damaging occurrences",
                    y="Losses", x="")

  
```

The single most economic damaging  event to crops was a **`r worst.crop.single.ev`, that occurred in `r worst.crop.single.st`, `r worst.crop.single.ct`, on `r worst.crop.single.dt`, causing U$ `r worst.crop.single.value` in losses.**


### Most Crop Damaging event in all time

Notice that are several occurrences of the same type of event along the time. 

Therefore, in order to know which is the worst type of event along all the years, we summed up the losses caused by each one of occurrences of this events. 

Notice that we are interested only in the worst of them, ie, the ones which are above the mean.


```{r crop-all, cache=TRUE}

# totals per event
crop.all.df <- crop.df %>% group_by(event) %>%
                summarise(total.raw = sum(crop.ev)) %>%
                arrange(desc(total.raw)) %>%
                mutate(media.raw = mean(total.raw), 
                        mediana.raw = median(total.raw),
                        total = dollar(total.raw),
                        mean = dollar(media.raw),
                        median = dollar(mediana.raw),
                        rank = seq_len(length(event))) %>%
                filter(total.raw > mean(total.raw))


# ordering events by rank, to plot later
crop.all.df$event <- factor(crop.all.df$event,
                levels = crop.all.df$event[order(crop.all.df$rank)])

# create color pallete for all events
colourCount.crop.all = length(unique(crop.all.df$event))
getPalette = colorRampPalette(brewer.pal(colourCount.crop.all, "Set1"))

# prepare text for inline R
worst.crop.all.ev <- crop.all.df$event[1]
worst.crop.total <- crop.all.df$total[1]

# a table
kable(crop.all.df[, c(8,1,5:7)], caption="Total losses by event")

```

```{r crop-all-plot, cache=TRUE, fig.cap="Total Crop Damages by event"}

# plt.crop.all <- ggplot(data=crop.all.df, aes(x=reorder(event, total.raw), y=total.raw, fill=event))

plt.crop.all <- ggplot(data=crop.all.df, aes(event, total.raw, fill=event))

plt.crop.all <- plt.crop.all + geom_bar(stat="identity") +
                
        geom_text(aes(label=ifelse(total.raw==max(total.raw),
                paste(event, dollar(max(total.raw)), sep=": "),'')),
                hjust=0,vjust=2) +
        geom_hline(aes(yintercept = media.raw), linetype=1) +
        # geom_hline(aes(yintercept = mediana.raw), linetype=2) +
        labs(title="All time", y="", x="") + 
                
        scale_y_continuous(labels = dollar)+
     
        theme(legend.position="none") +        
        scale_colour_manual(values = getPalette(colourCount.crop.all))+                
        theme(legend.title=element_blank()) +
        theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
        theme(plot.title = element_text(hjust = 0.5))                 

plt.crop.all + labs(title="Total Damages - All time",
             subtitle="Most crops damaging events of all time",
                     y="Losses", x="")

```



The most crop damaging event along the time is the **`r worst.crop.all.ev`. It has caused `r worst.crop.total` in losses.**


### Least crops damaging events

Just for curiosity, lets show now what are the less damaging among the  events:




```{r crop-less, cache=TRUE}

crop.all.df <- crop.df %>% group_by(event) %>%
                summarise(total.raw = sum(crop.ev)) %>%
                arrange(total.raw) %>%
                mutate(media.raw = mean(total.raw), 
                        mediana.raw = median(total.raw),
                        total = dollar(total.raw),
                        mean = dollar(media.raw),
                        median = dollar(mediana.raw),
                        rank = seq(length(event),1, by=-1))



kable(crop.all.df[1:10, c(8,1,5)], caption="Least crops damaging events")

```




