### Most injuring in a single occurrence
Most injuring in a single occurrence


In order to determine what were the most injuring events in a single occurrence, we need to see how injuries are distributed along the occurrences.



```{r injuring-distr-1, cache=TRUE}
rm(fatal.df,fatal.all.df, fatal95.df, qt) # cleannig house


injuring.df <- harm.df %>% filter(!is.na(injuries)) %>%
                select(1:6,8)

# quantiles
qt <- quantile(injuring.df$injuries, probs=seq(.975,1,0.002))

# distribution plot
plt.distr.inj0 <- ggplot(injuring.df, aes(injuries))

plt.distr.inj0 <- plt.distr.inj0 + geom_density(aes(y=..density..)) + xlim(0,0.5) + 
        labs(title="All events") +
        theme(plot.title = element_text(hjust = 0.5))


# display only the qts next to injuring events
qt[(length(qt)-(length(qt[qt>=1])+1)): length(qt)]
 
```

Looking at this distribution, we can infer that the vast majority of those occurrences were not injuring at all: **`r names(qt[qt==0])[length(qt[qt==0])]` occurrences didn't caused any injuries**

On the other hand, injuring occurrences had to have at least 1 injury.

Now, among the injuring occurrences, we are interested in the ones whose harm is beyond the confidence interval, ie. above 95% of the most common values. 

```{r injuring-distr-2, cache=TRUE}
# subset for harm events
injuring.df <- injuring.df %>% filter(injuries > 0) %>%
                arrange(desc(injuries)) %>%
                mutate(mean = mean(injuries), 
                median = median(injuries),
                rank = seq_len(length(event)))

# quantiles, same as 
# poisson.test(mean, conf.level = 0.95)

qt <- quantile(injuring.df$injuries, probs=seq(.975,1,0.005))
qt
```

Distribution plots


```{r injuring-distr-3, cache=TRUE}


# distribution plot
plt.distr.inj1 <- ggplot(injuring.df, aes(injuries))

plt.distr.inj1 <- plt.distr.inj1 + geom_density(aes(y=..density..)) + xlim(0,qt[1]) + 
        labs(title="Injuring events") +
        theme(plot.title = element_text(hjust = 0.5))       

grid.arrange(plt.distr.inj0, plt.distr.inj1, nrow=1, ncol=2, 
             bottom="Figure 2 - Distribution of Injuries: all events VS injuring events")
grid.rect(gp=gpar(fill=NA))

```

Looking at this distribution, we can infer that **95% of the injuring occurrences caused up to `r qt[1]` injuries**.

In this study, we looked on the 5% most injuring occurrences. 





```{r injuring-single-95, cache=TRUE}
# subset for 95% CI
injuring95.df <- filter(injuring.df, injuries>qt[1])

# create color pallete for all events
colourCount = length(unique(injuring95.df$event))
getPalette = colorRampPalette(brewer.pal(colourCount, "Set1"))

# print a table
kable(injuring95.df[,c(10,1:9)])

# prepare text for inline R
worst.injuring.single.ev <- injuring95.df$event[1]
worst.injuring.single.st <- injuring95.df$state[1]
worst.injuring.single.ct <- injuring95.df$countyname[1]
worst.injuring.single.dt <- injuring95.df$day[1]
worst.injuring.single.inj <- injuring95.df$injuries[1]
```


```{r injuring-single-plot, cache=TRUE}
# the plot
plt.inj.single <- ggplot(injuring95.df, aes(day, injuries, colour=event))

(plt.inj.single <- plt.inj.single + geom_point() +
        geom_text(aes(label=ifelse(rank <= 10,
                as.character(day),""),
                hjust=-.03,vjust=0.5)) +
        geom_hline(aes(yintercept = mean), linetype=2) +
        geom_hline(aes(yintercept = median), linetype=3) +
        labs(title="Most Injuring Events in a Single Occurrence",
                    y="Injuries", x="Ocurrence Date") +
                
        expand_limits(x=as.Date('2017-01-01'))+ #ok
        scale_colour_manual(values = getPalette(colourCount))+                
        theme(legend.title=element_blank()) +
        theme(legend.position="bottom") +
        guides(fill=guide_legend(nrow=5, byrow=TRUE)) +
        theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
        theme(plot.title = element_text(hjust = 0.5)) 

)  

```



The single most injuring event was a **`r worst.injuring.single.ev`, that occurred in `r worst.injuring.single.st`, `r worst.injuring.single.ct`, on `r worst.injuring.single.dt`, injuring `r worst.injuring.single.inj` people.**

However, if we compare this single awful event to the mean of injuries caused, we see that this is very unlikely to happen.



### Most injuring in all time
Most injuring in all time

Notice that are several occurrences of the same type of event along the time. 

Therefore, in order to know which is the worst type of event along all the years, we summed up the injuries caused by each one of occurrences of this events. 

Notice that we are interested only in the worst of them, ie, the ones which are above the mean.



```{r injuring-all, cache=TRUE}

# totals per event
injuring.all.df <- injuring.df %>% group_by(event) %>% 
        summarise(total = sum(injuries)) %>%
        arrange(desc(total)) %>% 
        mutate(mean = mean(total), median = median(total),
               rank = seq_len(length(event)))  %>%
        filter(total >= mean(total))

# create color pallete for all events
colourCount = length(unique(injuring.all.df$event))
getPalette = colorRampPalette(brewer.pal(colourCount, "Set1"))

# prepare text for inline R
worst.injuring.all.ev <- injuring.all.df$event[1]
worst.injuring.all.inj <- injuring.all.df$total[1]

# a table
kable(injuring.all.df[,c(5,1:4)])


```

The most injuring event along the time is the **`r worst.injuring.all.ev`. It has injuried `r worst.injuring.all.inj` people until now.**


```{r injuring-all-plot, cache=TRUE}

# the plot
plt.inj.all <- ggplot(data=injuring.all.df, aes(event, total, fill=event))

(plt.inj.all <- plt.inj.all + geom_bar(stat="identity") +
     
        geom_text(aes(label=ifelse(total==max(total),
                paste(event, max(total), sep=": "),'')),
                hjust=0,vjust=2) +
        geom_hline(aes(yintercept = mean), linetype=1) +
        geom_hline(aes(yintercept = median), linetype=2) +
        labs(title="Most Injuring events - all time", 
             y="Injuries", x="") + 
                
        theme(legend.position="none") +        
        scale_colour_manual(values = getPalette(colourCount))+                
        theme(legend.title=element_blank()) +
        theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
        theme(plot.title = element_text(hjust = 0.5))                 
                
)

```

Just for curiosity, lets show now what are the less injuring among the injuring events:

```{r injuring-less, cache=TRUE}

# sort for less dangerous, had to subset again due to previous filtering
injuring.all.df <- injuring.df %>% group_by(event) %>%
        summarise(total = sum(injuries)) %>% arrange(total) %>% 
        mutate(mean = mean(total), median = median(total),
               rank = seq(length(event),1, by=-1))

# a table
kable(injuring.all.df[1:10,c(5,1:2)])



```


