Unfortunatelly, these has been the worst counties for living in:


```{r cities.df, cache=TRUE}
# subset original data
cities.df <- harm.df %>% 
                filter(!(is.na(countyname) | is.na(state))) %>%
                select(6,5,7:14) %>%
                group_by(state, countyname) %>%
                summarise(fatalities = sum(fatalities, na.rm=TRUE),
                        injuries = sum(injuries, na.rm=TRUE),
                        prop.dmg = sum(prop.ev, na.rm=TRUE),
                        crop.dmg = sum(crop.ev, na.rm=TRUE)
                )

# city.fatal <- cities.df$countyname[which(cities.df$fatalities == max(cities.df$fatalities))]
# city.inj <- cities.df$countyname[which(cities.df$injuries == max(cities.df$injuries))]
# city.prop <- cities.df$countyname[which(cities.df$prop.ev == max(cities.df$prop.dmg))]
# city.crop <- cities.df$countyname[which(cities.df$crop.ev == max(cities.df$crop.dmg))]

```

### Worst fatality count
    

```{r fatal-city}

cities.fatal.df <- arrange(cities.df, desc(fatalities)) %>% ungroup(state, countyname) %>%
      mutate(rank = seq_len(length(fatalities)),
                prop.dmg = dollar(prop.dmg),
                crop.dmg = dollar(crop.dmg)
             )

kable(cities.fatal.df[1:10, c(7,1:6)])
worst.fatal.city.county <- cities.fatal.df$countyname[1]
worst.fatal.city.st <- cities.fatal.df$state[1]
worst.fatal.city.count <- cities.fatal.df$fatalities[1]




```


The county with the biggest fatality count is **`r worst.fatal.city.county`, in  `r worst.fatal.city.st`, with `r worst.fatal.city.count` people killed.**


### Worst injuries count
    

```{r inj-city}
rm(cities.fatal.df) # house cleanning
cities.inj.df <- arrange(cities.df, desc(injuries)) %>% ungroup(state, countyname) %>%
      mutate(rank = seq_len(length(injuries)),
                prop.dmg = dollar(prop.dmg),
                crop.dmg = dollar(crop.dmg)
             )

kable(cities.inj.df[1:10, c(7,1:6)])

worst.inj.city.county <- cities.inj.df$countyname[1]
worst.inj.city.st <- cities.inj.df$state[1]
worst.inj.city.count <- cities.inj.df$injuries[1]


```


The county with the biggest injuries count is **`r worst.inj.city.county`, in  `r worst.inj.city.st`, with `r worst.inj.city.count` people injuried.**


### Worst property losses



```{r prop-city}
rm(cities.inj.df) # house cleanning

cities.prop.df <- arrange(cities.df, desc(prop.dmg)) %>% ungroup(state, countyname) %>%
      mutate(rank = seq_len(length(prop.dmg)),
                prop.dmg = dollar(prop.dmg),
                crop.dmg = dollar(crop.dmg)
      )

kable(cities.prop.df[1:10, c(7,1:6)])

worst.prop.city.county <- cities.prop.df$countyname[1]
worst.prop.city.st <- cities.prop.df$state[1]
worst.prop.city.count <- cities.prop.df$prop.dmg[1]

```



The county with the biggest property losses is **`r worst.prop.city.county`, in  `r worst.prop.city.st`, with `r worst.prop.city.count` in losses.**

### Worst crops losses


```{r crop-city}
rm(cities.prop.df) # house cleanning


cities.crop.df <- arrange(cities.df, desc(crop.dmg)) %>% ungroup(state, countyname) %>%
      mutate(rank = seq_len(length(crop.dmg)),
                prop.dmg = dollar(prop.dmg),
                crop.dmg = dollar(crop.dmg)
      )

kable(cities.crop.df[1:10, c(7,1:6)])

worst.crop.city.county <- cities.crop.df$countyname[1]
worst.crop.city.st <- cities.crop.df$state[1]
worst.crop.city.count <- cities.crop.df$crop.dmg[1]

```

The county with the biggest croperty losses is **`r worst.crop.city.county`, in  `r worst.crop.city.st`, with `r worst.crop.city.count` in losses.**
