---
title: "NOAA Storm Database - worst cases"
author: "erickfis@gmail.com"
date: "22 de setembro de 2017"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Worst events of all time - fatalities


```{r fatal, cache=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(rmarkdown)
library(knitr)
library(RColorBrewer)
library(plotly)


load("data/harm.rda")
fatal.all.df <- harm.df %>% 
                filter(!is.na(fatalities)) %>%
                select(1:7) %>%
                filter(fatalities > 0) %>%
                arrange(desc(fatalities)) %>%
                mutate(mean = mean(fatalities), 
                median = median(fatalities),
                rank = seq_len(length(event))) %>% 
        group_by(event) %>%
        summarise(total = sum(fatalities)) %>% arrange(desc(total)) %>% 
        mutate(mean = mean(total), median = median(total),
               rank = seq_len(length(event))) %>%
        filter(total > mean(total))

# ordering events by rank, to plot later
fatal.all.df$event <- factor(fatal.all.df$event,
                levels = fatal.all.df$event[order(fatal.all.df$rank)])


# create color pallete for all events
colourCount.fatal.all = length(unique(fatal.all.df$event))
getPalette = colorRampPalette(brewer.pal(colourCount.fatal.all, "Set1"))


# the plot
plt.fatal.all <- ggplot(data=fatal.all.df, aes(event, total, fill=event,
                        text = paste(event, "<br>", total, "fatalities")))



plt.fatal.all <- plt.fatal.all + geom_bar(stat="identity") +
        geom_hline(aes(yintercept = mean, 
                       text=paste("mean", round(mean,2))), linetype=1) +
        labs(title="All time", y="", x="") + 
        theme(legend.position="none") +        
        scale_colour_manual(values = getPalette(colourCount.fatal.all))+                theme(legend.title=element_blank()) +
        theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
        theme(plot.title = element_text(hjust = 0.5))                 

ggplotly(plt.fatal.all + labs(title="Total Fatalities - All time",
             subtitle="Most fatal events of all time",
                     y="Fatalities", x=""), tooltip = "text")


```


## Source

The U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database tracks characteristics of major storms and weather events in the United States, including when and where they occur, as well as estimates of any fatalities, injuries, and property damage.

The database currently contains data from January 1950 to January 2017, as entered by NOAA's National Weather Service (NWS).

The database can be found on:

https://www.ncdc.noaa.gov/stormevents/ftp.jsp
