# Data Processing 

In order to answer our questions, the original database needed to be treated from its raw form to a more useful format.


```{r libraries}

library(scales)
library(stringr)
library(data.table)
library(chron)
library(dplyr)
library(lubridate)
library(ggplot2)
library(rmarkdown)
library(RColorBrewer)
library(gridExtra)
library(grid)
library(pander)

```



```{r inittial-load, cache=TRUE}
# Reading original database:

#fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"

#download.file(fileUrl, "StormData.bz2", method = "curl")

# looks for the already treated data base harm.rds. If not present, runs from the original
# data base. Else, just loads harm.rds

arquivos <- dir()
arquivos <- arquivos[grep("harm", arquivos)]


if (arquivos!="harm.rds"){

#Full data
dados <- fread(sprintf("bzcat %s | tr -d '\\000'", "StormData.bz2"), na.strings = "")
dados <-tbl_df(dados)

# this do a sample data base, with 50000 obs, used for speeding up initial works:
# linhas <- nrow(dados)
# linhas <- sample(linhas,50000)
# dataS <- dados[linhas,]
# write.csv(dataS, "StormData")

# dados <- fread(sprintf("bzcat %s | tr -d '\\000'", "StormData-sample.bz2"))
# dados <-tbl_df(dados)
# dados <- select(dados, -1)



# treating var names
names(dados) <- gsub("_", ".", tolower(names(dados)))


# select desired vars
harm.df <- dados %>% select(evtype, mag, state, countyname, bgn.date, end.date, 23:28)

# treat vars
harm.df <- harm.df %>% 
        mutate(bgn.date = mdy_hms(bgn.date), end.date = mdy_hms(end.date),  
               day = as.Date(bgn.date, "%m/%d/%Y"), 
               duration = -as.period(interval(end.date, bgn.date)),
               event = tolower(as.character(evtype)),
               countyname =strtrim(countyname,9)) %>%
        select(event, 2, day, duration, 3:4, 7:12)

# fixing exp for economic data

unique(harm.df$propdmgexp) # this shows what we need to look for

harm.df$propdmgexp[is.na(harm.df$propdmgexp)] <- as.character(0)
harm.df$propdmgexp[which(harm.df$propdmgexp %in% c("+","?", "-"))] <- as.character(0)

harm.df$propdmgexp[which(harm.df$propdmgexp %in% c("H","h"))] <- as.character(2)
harm.df$propdmgexp[which(harm.df$propdmgexp %in% c("K","k"))] <- as.character(3)
harm.df$propdmgexp[which(harm.df$propdmgexp %in% c("M","m"))] <- as.character(6)
harm.df$propdmgexp[which(harm.df$propdmgexp %in% c("B","b"))] <- as.character(9)
harm.df$propdmgexp <- as.numeric(harm.df$propdmgexp)


unique(harm.df$cropdmgexp) # this shows what we need to look for

harm.df$cropdmgexp[is.na(harm.df$cropdmgexp)] <- as.character(0)
harm.df$cropdmgexp[which(harm.df$cropdmgexp == "?")] <- as.character(0)

harm.df$cropdmgexp[which(harm.df$cropdmgexp %in% c("K","k"))] <- as.character(3)
harm.df$cropdmgexp[which(harm.df$cropdmgexp %in% c("M","m"))] <- as.character(6)
harm.df$cropdmgexp[which(harm.df$cropdmgexp %in% c("B","b"))] <- as.character(9)
harm.df$cropdmgexp <- as.numeric(harm.df$cropdmgexp)

harm.df <- mutate(harm.df, prop.ev = propdmg*10^propdmgexp,
                crop.ev = cropdmg*10^cropdmgexp)


# treating event types

eventos <- harm.df$event
# first, need to see what are the event types
contagem <- sort(table(eventos))

# them we create this list of terms
 
lista.search <- c(
"dry",        
"fog",        
"wind",
"winter",
"slide",
"snow",
"flood",
"fld",
"cold|freez",
"hurricane",
"tornado",
"rain|precip",
"hail",
"heat|warm",
"tide",
"storm",
"record",
"blizzard",
"fire",
"funnel",
"surf")

lista.replace <- c(
"drought",
"fog",        
"wind",
"winter",
"slide",
"snow",
"flood",
"flood",
"cold",
"hurricane",
"tornado",
"rain",
"hail",
"heat",
"tide",
"storm",
"record temperature",
"blizzard",
"fire",
"funnel",
"surf")


for (i in 1:length(lista.search)) {
        eventos[grepl(lista.search[i], eventos)] <- lista.replace[i]
        
}

# lets group the events whose count is < 5 and call it "other"
contagem <- sort(table(eventos))
outros <- names(contagem[contagem<5])
eventos[eventos %in% outros] <- "other"
# sort(table(eventos))

# returning treated events
harm.df$event <- toupper(eventos)


# Treating County Names
cidades <- toupper(harm.df$countyname)
cidades <- str_trim(cidades)

cidades <- gsub("_| |-", ".", cidades)

cidades[grep("\\.\\.$", cidades)] <- gsub("\\.\\.", "",
                                cidades[grep("\\.\\.$", cidades)]
                                        )

cidades <- sapply(strsplit(cidades, split=">", fixed=TRUE), function(x) (x[1]))
cidades <- sapply(strsplit(cidades, split="(", fixed=TRUE), function(x) (x[1]))
cidades <- sapply(strsplit(cidades, split=",", fixed=TRUE), function(x) (x[1]))

harm.df$countyname <- cidades

rm(dados, cidades) # house cleanning
saveRDS(harm.df, "harm.rds")

} else {
        harm.df <- readRDS("harm.rds")
}



```

The necessary transformations were:

- sanitized var names
- evaluated duration of events
- evaluated damages values according to multipliers provided
- sanitized and grouped similar events: strong snow, heavy snow and light snow all became just "snow"
- sanitized county names


This database has `r nrow(harm.df)` observations. Each observation corresponds to an event occurrence.

To determine the most harmful events to human health, we checked the variables related to human health, which are "fatalities" and "injuries".

To determine the most harmful events to economy, we checked the variables related to economic measures, from "propdmg" through "cropdmgexp". 

Also, in order to analyse various occurrences of the same event, we measured the duration of the event, its magnitude and where the event occurred (state and county name).


This is a really big database whose data has been being registered by a lot of different people since 1950. Thus, as expected, there are variations on how people registered events. 

For example, the string "snow" was used to register a lot of events. They are the same type of event, but count as different:

This is why we decided to filter those events: we grouped them by its common strings.


