---
title: "NOAA Storm Database - worst cases"
author: "erickfis"
date: "`r format(Sys.time(), '%Y %B, %d')`"
output:
  md_document

# output: 
#         html_document:
#                 toc: true
#                 toc_depth: 3
#                 number_sections: true
#                 self_contained: yes
#                 # fig_width: 11
#                 # fig_height: 7
#                 fig_crop: false
#         classoption: landscape



# output:
#   pdf_document:
#         toc: true
#         toc_depth: 3
#         number_sections: true
#         keep_tex: true

# output:
#         odt_document:
#         toc: true
#         toc_depth: 3
#         number_sections: true

---

```{r knit-rules, echo=FALSE}
library(knitr)

opts_chunk$set(echo = TRUE, message = FALSE, include = TRUE, 
               warning = FALSE, autodep = TRUE, options(scipen=999))

```

# NOAA Storm Database - worst cases


In this study we have analysed the NOAA Storm Database in order to determine what are the worst natural catastrophic events, both in terms of public health and in economic impact.

The U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database tracks characteristics of major storms and weather events in the United States, including when and where they occur, as well as estimates of any fatalities, injuries, and property damage.

The database currently contains data from January 1950 to January 2017, as entered by NOAA's National Weather Service (NWS).

The database can be found on:

https://www.ncdc.noaa.gov/stormevents/ftp.jsp


RPubs version, with fewer plots, for Coursera:
http://rpubs.com/erickfis/noaa


# Objective


The goal of this study is to answer the questions:

1. Across the United States, which types of events were the most harmful with respect to population health ever recorded in a single occurrence?

1. Which types of events caused most harm to population health along all those years?

1. Which types of events had the greatest economic consequences in a single occurrence?

1. Which types of events had the greatest economic consequences along all those years?

1. Which were the places that were subject to the greatest losses, both in terms of human health and economic losses.

# Data Processing 

```{r child = 'processamento.Rmd'}
```


# Human health: the most harmfull events

We have determined what events did more harm to human health.

There were occurrences that caused zero fatalities but a lot of injuries.  The inverse is also true, so we did a separate analysis to fatal and non-fatal events.


## Fatal Occurrences

```{r child = 'fatal.Rmd'}
```

## Injuring Occurrences

```{r child = 'injuring.Rmd'}
```


# Economy: the the most harmfull events

We have determined what events did more harm to economy, both in terms of property and crops damage.

There were events that causes zero property damage but a lot of crop damage. The inverse is also true, so we did a separate analysis to property VS crop damaging events.

## Property losses

```{r child = 'prop.Rmd'}
```

## Crop losses

```{r child = 'crop.Rmd'}
```



# Most aflicted locations

Most afflicted locations

We have determined what locations had the worst outcome from those events, both in terms of human health and economic losses.


```{r child = 'locations.Rmd'}
```


# Results

```{r child = 'resultados.Rmd'}
```






