---
author: "erickfis"
date: "27 de abril de 2017"
output: 
  md_document
---

```{r knit-rules, echo=FALSE}
library(knitr)

opts_chunk$set(echo = TRUE, message = FALSE, include = TRUE, 
               warning = FALSE, autodep = TRUE, options(scipen=999))

```



# NOAA Storm Database: the worst events

NOAA Storm Database Study: the worst events

[HTML version](https://github.com/erickfis/NOAA-Storm-Database/blob/master/NOAA-stormDB.html)


In this study we will analise the NOAA Storm Database in order to determine what are the worst natural catastrofic events, both in terms of public health and in economic impact.

The U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database tracks characteristics of major storms and weather events in the United States, including when and where they occur, as well as estimates of any fatalities, injuries, and property damage.

The database currently contains data from January 1950 to January 2017, as entered by NOAA's National Weather Service (NWS).

The database can be found on:

https://www.ncdc.noaa.gov/stormevents/ftp.jsp

## Data Processing 

```{r child = 'processamento.Rmd'}
```


## Human health: the most harmfull events

There are occurrences that caused zero fatalities but a lot of injuries, so we did a separate analisys to fatal and non-fatal events.


### Fatal Occurrences

```{r child = 'fatal.Rmd'}
```

### Injuring Occurrences

```{r child = 'injuring.Rmd'}
```


## Economy: the the most harmfull events

Here we determine what are events did more harm to economy, both int terms of property and crops damage.

There are events that causes zero property damage but a lot of crop damage. The inverse is also true, so we did a separate analisys to property VS crop damaging events.

### Property losses

```{r child = 'prop.Rmd'}
```

### Crop losses

```{r child = 'crop.Rmd'}
```

## Results


```{r child = 'resultados.Rmd'}
```

