
## Population Health


```{r health-plot, fig.width=10, fig.height=10}

# plist <- list(plt.fatal.single, plt.fatal.all, plt.inj.single, plt.inj.all)
# n <- length(plist)
# nCol <- floor(sqrt(n))
# do.call("grid.arrange", c(plist, ncol=nCol))

grid.arrange(plt.fatal.single, plt.fatal.all, plt.inj.single, plt.inj.all, 
             nrow=2, ncol=2,
             bottom="Health impact")
grid.rect(gp=gpar(fill=NA))
```



The single most fatal event was a **`r worst.fatal.single.ev`, that occurred in `r worst.fatal.single.st`, `r worst.fatal.single.ct`, on `r worst.fatal.single.dt`, killing `r worst.fatal.single.kill` people.**


The most fatal event along the time is the **`r worst.fatal.all.ev`. It has killed `r worst.fatal.all.kill` people until now.**


The single most injuring event was a **`r worst.injuring.single.ev`, that occurred in `r worst.injuring.single.st`, `r worst.injuring.single.ct`, on `r worst.injuring.single.dt`, injuring `r worst.injuring.single.inj` people.**

The most injuring event along the time is the **`r worst.injuring.all.ev`. It has injuried `r worst.injuring.all.inj` people until now.**

## Economic Damages


```{r economic-plot, fig.width=10, fig.height=10}
grid.arrange(plt.prop.single, plt.prop.all,
             plt.crop.single, plt.crop.all,
             nrow=2, ncol=2,
             bottom="Economic impact")
grid.rect(gp=gpar(fill=NA))
```


The single most economic damaging  event to properties was a **`r worst.prop.single.ev`, that occurred in `r worst.prop.single.st`, `r worst.prop.single.ct`, on `r worst.prop.single.dt`, causing U$ `r worst.prop.single.value` in losses**.


The most property damaging event along the time is the **`r worst.prop.all.ev`. It has caused `r worst.prop.total` in losses.**


The single most economic damaging  event to crops was a **`r worst.crop.single.ev`, that occurred in `r worst.crop.single.st`, `r worst.crop.single.ct`, on `r worst.crop.single.dt`, causing U$ `r worst.crop.single.value` in losses**.


The most crop damaging event along the time is the **`r worst.crop.all.ev`. It has caused `r worst.crop.total` in losses.**


## Most aflicted locations

The county with the biggest fatality count is **`r worst.fatal.city.county`, in  `r worst.fatal.city.st`, with `r worst.fatal.city.count` people killed.**


The county with the biggest injuries count is **`r worst.inj.city.county`, in  `r worst.inj.city.st`, with `r worst.inj.city.count` people injuried.**

The county with the biggest property losses is **`r worst.prop.city.county`, in  `r worst.prop.city.st`, with `r worst.prop.city.count` in losses.**

The county with the biggest croperty losses is **`r worst.crop.city.county`, in  `r worst.crop.city.st`, with `r worst.crop.city.count` in losses.**


## Distribution of data

```{r distribution, fig.width=10, fig.height=3}
grid.arrange(plt.distr.fatal0, plt.distr.fatal1, plt.distr.inj0, plt.distr.inj1, 
             nrow=1, ncol=4, 
             bottom="Population distribution - fatalities and injuries")
grid.rect(gp=gpar(fill=NA))             
             
             
grid.arrange(plt.distr.prop0, plt.distr.prop1, plt.distr.crop0, plt.distr.crop1, 
             nrow=1, ncol=4, 
             bottom="Population distribution - property and crops losses")
grid.rect(gp=gpar(fill=NA))   
```
