---
title       : NOAA - Worst natural disasters
subtitle    : Most damaging types of natural disasters
date: "28 de setembro de 2017"
output: ioslides_presentation
mode        : standalone # {standalone, draft}

---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```


## NOAA Storm Database

### Worst Weather Events

The U.S. National Oceanic and Atmospheric Administration's (NOAA) storm database tracks characteristics of major storms and weather events in the United States, including when and where they occur, as well as estimates of any fatalities, injuries, and property damage.

The database currently contains data from January 1950 to January 2017, as entered by NOAA's National Weather Service (NWS).

Full study: https://erickfis.github.io/NOAA-Storm-Database/

erickfis@gmail.com

2017-09-28



```{r libs, message=FALSE}
# library(shiny)
# library(plotly)
library(dplyr)
library(lubridate)
library(ggplot2)
library(rmarkdown)
library(knitr)
library(scales)
library(rCharts)
library(RColorBrewer)
```


```{r economic-db, message=FALSE, cache=TRUE}
load("../data/harm.rda")

Property <- harm.df %>% filter(!is.na(prop.ev)) %>% select(1:6,9) %>% 
                filter(prop.ev > 0) %>%
                group_by(event) %>%
                summarise(total.raw = sum(prop.ev)) %>%
                arrange(desc(total.raw)) %>%
                mutate(
                        media.raw = mean(total.raw),
                        rank = seq_len(length(event))
                       ) %>%
                filter(total.raw > mean(total.raw))

Property$event <- factor(Property$event,
                           levels = Property$event[order(Property$rank)])



Crops <- harm.df %>% filter(!is.na(crop.ev)) %>% select(1:6,10) %>%
                filter(crop.ev > 0) %>%
                group_by(event) %>%
                # summarise(total.raw = sum(crop.ev)) %>%
                summarise(total.raw = sum(crop.ev)) %>%
                arrange(desc(total.raw)) %>%
                mutate(
                        media.raw = mean(total.raw),
                        rank = seq_len(length(event))
                        ) %>%
                filter(total.raw > mean(total.raw))

Crops$event <- factor(Crops$event,
                           levels = Crops$event[order(Crops$rank)])

```


## Property losses, in billions $

```{r props, results='asis'}


colourCount = length(unique(Property$event))
colors <- brewer.pal(colourCount, "Set1")
Property$colors <- colors
Property$total.raw <- Property$total.raw/1000000000


p2 <-  nPlot(total.raw ~ event, data = Property, type = 'discreteBarChart')
p2$chart(color = Property$colors)
# p2$print(include_assets=T)
p2$show('inline', include_assets = TRUE, cdn = TRUE)

```


## Crops losses, in billions $

```{r crops, results='asis'}

colourCount = length(unique(Crops$event))
colors <- brewer.pal(colourCount, "Set1")
Crops$colors <- colors
Crops$total.raw <- Crops$total.raw/1000000000


p2 <-  nPlot(total.raw ~ event, data = Crops, type = 'discreteBarChart')
p2$chart(color = Crops$colors)

p2$show('inline', include_assets = TRUE, cdn = TRUE)
```


## Worst Weather Events - Fatalities


```{r health-db, message=FALSE, cache=TRUE}

Fatalities <- harm.df %>%
        filter(!is.na(fatalities)) %>% select(1:7) %>%
        filter(fatalities > 0) %>%
        group_by(event) %>%
        summarise(total = sum(fatalities)) %>% arrange(desc(total)) %>%
        mutate(mean = mean(total), median = median(total),
               rank = seq_len(length(event))
               ) %>%
        filter(total > mean(total))


# ordering events by rank, to plot later
Fatalities$event <- factor(Fatalities$event,
                              levels = Fatalities$event[order(Fatalities$rank)])


Injuries <- harm.df %>% filter(!is.na(injuries)) %>% select(1:6,8) %>%
                filter(injuries > 0) %>%
                group_by(event) %>%
                summarise(total = sum(injuries)) %>%
                arrange(desc(total)) %>%
                mutate(mean = mean(total), median = median(total),
                        rank = seq_len(length(event))
                       )  %>%
                filter(total >= mean(total))

Injuries$event <- factor(Injuries$event,
                           levels = Injuries$event[order(Injuries$rank)])

```

```{r fatalities-plot, results='asis'}

colourCount = length(unique(Fatalities$event))
colors <- brewer.pal(colourCount, "Set1")
Fatalities$colors <- colors

p2 <-  nPlot(total ~ event, data = Fatalities, type = 'discreteBarChart')
p2$chart(color = Fatalities$colors)

# p2$print(include_assets=T)
p2$show('inline', include_assets = TRUE, cdn = TRUE)
       
```

## Worst Weather Events - Injuries

```{r injuries-plot, results='asis'}

colourCount = length(unique(Injuries$event))
colors <- brewer.pal(colourCount, "Set1")
Injuries$colors <- colors

p2 <-  nPlot(total ~ event, data = Injuries, type = 'discreteBarChart')
p2$chart(color = Injuries$colors)

# p2$print(include_assets=T)
p2$show('inline', include_assets = TRUE, cdn = TRUE)
        
```


