## Injuring Occurrences


### Most injuring in a single occurrence
Most injuring in a single occurrence


In order to determine what were the most injuring events in a single occurrence, we need to see how injuries are distributed along the occurrences.



```{r injuring-distr-1, cache=TRUE}
rm(fatal.df,fatal.all.df, fatal95.df, qt) # cleannig house


injuring.df <- harm.df %>% filter(!is.na(injuries)) %>%
                select(1:6,8)

# quantiles
qt <- quantile(injuring.df$injuries, probs=seq(.975,1,0.002))

# distribution plot
plt.distr.inj0 <- ggplot(injuring.df, aes(injuries))

plt.distr.inj0 <- plt.distr.inj0 + geom_density(aes(y=..scaled..)) + xlim(0,0.5) + 
        labs(title="All events") +
        theme(plot.title = element_text(hjust = 0.5))


# display only the qts next to injuring events
#qt[(length(qt)-(length(qt[qt>=1])+1)): length(qt)]
 
```

Looking at this distribution, we can infer that the vast majority of those occurrences were not injuring at all: **`r names(qt[qt==0])[length(qt[qt==0])]` occurrences didn't caused any injuries**

On the other hand, injuring occurrences had to have at least 1 injury.

Now, among the injuring occurrences, we are interested in the ones whose harm is beyond the confidence interval, ie. above 99% of the most common values. 

```{r injuring-distr-2, cache=TRUE}
# subset for harm events
injuring.df <- injuring.df %>% filter(injuries > 0) %>%
                arrange(desc(injuries)) %>%
                mutate(mean = mean(injuries), 
                median = median(injuries),
                rank = seq_len(length(event)))


qt <- quantile(injuring.df$injuries, probs=seq(.999,1,0.005))
#qt
```

Looking at this distribution, we can infer that **99.8% of the injuring occurrences caused up to `r qt[1]` injuries**.



```{r injuring-distr-3, cache=TRUE}


# distribution plot
plt.distr.inj1 <- ggplot(injuring.df, aes(injuries))

plt.distr.inj1 <- plt.distr.inj1 + geom_density(aes(y=..scaled..)) + xlim(0,(qt[1]/20)) + 
        labs(title="Injuring events") +
        theme(plot.title = element_text(hjust = 0.5))       

```


```{r inj-distribution, fig.width=7, fig.height=3, fig.cap="Population distribution for Injuries / occurrences"}
grid.arrange(plt.distr.inj0, plt.distr.inj1,
             nrow=1, ncol=2)
grid.rect(gp=gpar(fill=NA))
```


In this study, we looked on the 1% most injuring occurrences. 





```{r injuring-single-95, cache=TRUE}
# subset for 99% ocurrences
injuring95.df <- filter(injuring.df, injuries>qt[1])

# create color pallete for all events
colourCount.inj.single = length(unique(injuring95.df$event))
getPalette = colorRampPalette(brewer.pal(colourCount.inj.single, "Set1"))


# prepare text for inline R
worst.injuring.single.ev <- injuring95.df$event[1]
worst.injuring.single.st <- injuring95.df$state[1]
worst.injuring.single.ct <- injuring95.df$county[1]
worst.injuring.single.dt <- injuring95.df$day[1]
worst.injuring.single.inj <- injuring95.df$injuries[1]
worst.injuring.single.mean <- round(injuring95.df$mean[1],2)
worst.injuring.single.median <- round(injuring95.df$median[1],2)

# print a table

kable(injuring95.df[, c(10,1,3,5:7)], 
      caption=paste("Worst injuring occurrences, mean = ",
                    worst.injuring.single.mean, " and median = ",
                    worst.injuring.single.median))
```


```{r injuring-single-plot, cache=TRUE, fig.cap="Worst injuring occurrences"}
# the plot
plt.inj.single <- ggplot(injuring95.df, aes(day, injuries, colour=event))

plt.inj.single <- plt.inj.single + geom_point() +
        geom_text(aes(label=ifelse(rank == 1,
                 paste0(as.character(day), ": ", injuries, " injuried") ,""),
                hjust=.5,vjust=1.5)) +
        geom_text(aes(label=ifelse(rank %in% c(2,3),
                 paste0(as.character(day), ": ", injuries, " injuried") ,""),
                hjust=-.03,vjust=0.5)) +
        # geom_hline(aes(yintercept = mean), linetype=2) +
        # geom_hline(aes(yintercept = median), linetype=3) +
        labs(title="Most Injuring",
                    y="", x="") +
                
        # expand_limits(x=as.Date('2017-01-01'))+ #ok
        scale_colour_manual(values = getPalette(colourCount.inj.single))+                
        theme(legend.title=element_blank()) +
        theme(legend.position="bottom") +
        guides(fill=guide_legend(nrow=5, byrow=TRUE)) +
        theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
        theme(plot.title = element_text(hjust = 0.5)) 


        # scale_y_continuous(expand = c(0, 0)) +
        # geom_hline(aes(yintercept = mean), linetype=2) +
        # geom_hline(aes(yintercept = median), linetype=3)
plt.inj.single + labs(title="Worst injuring occurrences",
                    y="Injuries", x="") 


```



The single most injuring event was a **`r worst.injuring.single.ev`, that occurred in `r worst.injuring.single.st`, `r worst.injuring.single.ct`, on `r worst.injuring.single.dt`, injuring `r worst.injuring.single.inj` people.**

However, if we compare this single awful event to the mean of injuries caused, we see that this is very unlikely to happen.



### Most injuring in all time

Notice that are several occurrences of the same type of event along the time. 

Therefore, in order to know which is the worst type of event along all the years, we summed up the injuries caused by each one of occurrences of this events. 

Notice that we are interested only in the worst of them, ie, the ones which are above the mean.



```{r injuring-all, cache=TRUE}

# totals per event
injuring.all.df <- injuring.df %>% group_by(event) %>% 
        summarise(total = sum(injuries)) %>%
        arrange(desc(total)) %>% 
        mutate(mean = mean(total), median = median(total),
               rank = seq_len(length(event)))  %>%
        filter(total >= mean(total))

# ordering events by rank, to plot later
injuring.all.df$event <- factor(injuring.all.df$event,
                levels = injuring.all.df$event[order(injuring.all.df$rank)])


# create color pallete for all events
colourCount.inj.all = length(unique(injuring.all.df$event))
getPalette = colorRampPalette(brewer.pal(colourCount.inj.all, "Set1"))

# prepare text for inline R
worst.injuring.all.ev <- injuring.all.df$event[1]
worst.injuring.all.inj <- injuring.all.df$total[1]
worst.injuring.all.mean <- round(injuring.all.df$mean[1],2)
worst.injuring.all.median <- round(injuring.all.df$median[1],2)

# a table
kable(injuring.all.df[,c(5,1:2)],
      caption=paste("Total injuries by event, mean = ",
                    worst.injuring.all.mean, " and median = ",
                    worst.injuring.all.median))


```

The most injuring event along the time is the **`r worst.injuring.all.ev`. It has injuried `r worst.injuring.all.inj` people until now.**


```{r injuring-all-plot, fig.cap="Total Injuries by event"}

# the plot
plt.inj.all <- ggplot(data=injuring.all.df, aes(event, total, fill=event))

plt.inj.all <- plt.inj.all + geom_bar(stat="identity") +
        geom_text(aes(label=ifelse(total==max(total),
                paste0(event, ": ", max(total), " injuried"),'')),
                hjust=0,vjust=2) +
        geom_hline(aes(yintercept = mean), linetype=1) +
        # geom_hline(aes(yintercept = median), linetype=2) +
        labs(title="All time", 
             y="", x="") + 
                
        theme(legend.position="none") +        
        scale_colour_manual(values = getPalette(colourCount.inj.all))+                
        theme(legend.title=element_blank()) +
        theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
        theme(plot.title = element_text(hjust = 0.5))                 
                

plt.inj.all  + labs(title="Total Injuries - All time",
             subtitle="Most injuring events of all time",
                     y="Injuries", x="")

```

### Least injuring events



Just for curiosity, lets show now what are the less injuring among the injuring events:

```{r injuring-less, cache=TRUE}

# sort for less dangerous, had to subset again due to previous filtering
injuring.all.df <- injuring.df %>% group_by(event) %>%
        summarise(total = sum(injuries)) %>% arrange(total) %>% 
        mutate(mean = mean(total), median = median(total),
               rank = seq(length(event),1, by=-1))

# a table
kable(injuring.all.df[1:10,c(5,1:2)], caption="Least injuring events")

#kable(injuring.all.df[1:10,c(5,1:2)])


```


