# Most aflicted locations


We have determined what locations had the worst outcome from those events, both in terms of human health and economic losses.



Unfortunately, these has been the worst counties for living in:


```{r cities.df, cache=TRUE}
rm(crop.df, crop.all.df, crop95.df) # house cleanning

# subset original data
cities.df <- harm.df %>% 
                filter(!(is.na(county) | is.na(state))) %>%
                select(6,5,7:10) %>%
                group_by(state, county) %>%
                summarise(fatalities = sum(fatalities, na.rm=TRUE),
                        injuries = sum(injuries, na.rm=TRUE),
                        prop.dmg = sum(prop.ev, na.rm=TRUE),
                        crop.dmg = sum(crop.ev, na.rm=TRUE)
                )
```


## Worst fatality count
    

```{r fatal-city}

cities.fatal.df <- arrange(cities.df, desc(fatalities)) %>% ungroup(state, county) %>%
      mutate(rank = seq_len(length(fatalities)),
                prop.dmg = dollar(prop.dmg),
                crop.dmg = dollar(crop.dmg)
             )

kable(cities.fatal.df[1:10, c(7,1:6)], caption="Total fatalities by county")
#kable(cities.fatal.df[1:10, c(7,1:6)])

worst.fatal.city.county <- cities.fatal.df$county[1]
worst.fatal.city.st <- cities.fatal.df$state[1]
worst.fatal.city.count <- cities.fatal.df$fatalities[1]
```


The county with the biggest fatality count is **`r worst.fatal.city.county`, in  `r worst.fatal.city.st`, with `r worst.fatal.city.count` people killed.**

\newpage

## Worst injuries count
    

```{r inj-city}
rm(cities.fatal.df) # house cleanning
cities.inj.df <- arrange(cities.df, desc(injuries)) %>% ungroup(state, county) %>%
      mutate(rank = seq_len(length(injuries)),
                prop.dmg = dollar(prop.dmg),
                crop.dmg = dollar(crop.dmg)
             )

kable(cities.inj.df[1:10, c(7,1:6)], caption="Total injuries by county")
#kable(cities.inj.df[1:10, c(7,1:6)])

worst.inj.city.county <- cities.inj.df$county[1]
worst.inj.city.st <- cities.inj.df$state[1]
worst.inj.city.count <- cities.inj.df$injuries[1]


```


The county with the biggest injuries count is **`r worst.inj.city.county`, in  `r worst.inj.city.st`, with `r worst.inj.city.count` people injuried.**


## Worst property losses



```{r prop-city}
rm(cities.inj.df) # house cleanning

cities.prop.df <- arrange(cities.df, desc(prop.dmg)) %>% ungroup(state, county) %>%
      mutate(rank = seq_len(length(prop.dmg)),
                prop.dmg = dollar(prop.dmg),
                crop.dmg = dollar(crop.dmg)
      )

kable(cities.prop.df[1:10, c(7,1:6)], caption="Total property losses by county")
#kable(cities.prop.df[1:10, c(7,1:6)])

worst.prop.city.county <- cities.prop.df$county[1]
worst.prop.city.st <- cities.prop.df$state[1]
worst.prop.city.count <- cities.prop.df$prop.dmg[1]

```



The county with the biggest property losses is **`r worst.prop.city.county`, in  `r worst.prop.city.st`, with `r worst.prop.city.count` in losses.**



## Worst crops losses


```{r crop-city}
rm(cities.prop.df) # house cleanning


cities.crop.df <- arrange(cities.df, desc(crop.dmg)) %>% ungroup(state, county) %>%
      mutate(rank = seq_len(length(crop.dmg)),
                prop.dmg = dollar(prop.dmg),
                crop.dmg = dollar(crop.dmg)
      )

kable(cities.crop.df[1:10, c(7,1:6)], caption="Total crops losses by county")
#kable(cities.crop.df[1:10, c(7,1:6)])


worst.crop.city.county <- cities.crop.df$county[1]
worst.crop.city.st <- cities.crop.df$state[1]
worst.crop.city.count <- cities.crop.df$crop.dmg[1]

```

The county with the biggest croperty losses is **`r worst.crop.city.county`, in  `r worst.crop.city.st`, with `r worst.crop.city.count` in losses.**
